%%
%% This is file `paresse-en.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% paresse.dtx  (with options: `doc,en')
%% This is a generated file.
%% 
%% Copyright (C) 2003-2010 Y. Henel, le TeXnicien de surface
%% <le.texnicien.de.surface@wanadoo.fr>
%% 
%% This file may be distributed and/or modified under the conditions of
%% the LaTeX Project Public License, either version 1.2 of this license
%% or (at your option) any later version.  The latest version of this
%% license is in:
%% 
%%    http://www.latex-project.org/lppl.txt
%% 
%% and version 1.2 or later is part of all distributions of LaTeX version
%% 1999/12/01 or later.
\def\fileversion{v3}
\def\filedate{2010/03/21}
\documentclass[a4paper]{ltxdoc}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[tame]{paresse}
\usepackage{xspace,url,array,ifthen}
\IfFileExists{lmodern.sty}{\usepackage{lmodern}}{}
\usepackage[frenchb,english]{babel}
\usepackage[colorlinks=true,
            linkcolor=blue,
            urlcolor=blue,
            citecolor=blue]{hyperref}
\makeatletter
\newcommand\DescribeOption[1]{\@bsphack \marginpar{\raggedleft
    \textsc{#1}}\index{#1 (option)}\@esphack}
\newcommand\BOP{\discretionary{}{}{}}
\newcommand\Option[1]{\textsc{#1}}
\newcommand\MODULE[1]{\textsf{#1}}
\newcommand\NOUVO{%
  \marginpar{\hspace*{\stretch{1}}%
    \textbf{N (v3)}\hspace*{.5em}}\xspace}

\begin{document}
\GetFileInfo{paresse.sty} % marche pô! :(

\title{The package \MODULE{paresse}\thanks{This document corresponds
    to the file \MODULE{\filename}~\fileversion, dated \filedate.}
  \mbox{}\thanks{English translation by the author. Any comment
    about the translation is welcome.}}
\author{Le \TeX nicien de surface\\
\url{le.texnicien.de.surface@wanadoo.fr}}
\maketitle{}

\begin{abstract}
  This package implements an example from T.~\textsc{Lachand-Robert}
  in~\cite{tlachand}. It provides a means of typing isolated greek
  letters with the character § activated and redefined. Instead of
  |\(\alpha\)| one types |§a| to obtain \(\alpha\).

  The code is not commented in English any more. Sorry. See the French
  documentation for French commented code in \texttt{paresse.pdf}.
\end{abstract}
\begin{otherlanguage}{french}
  \begin{abstract}
    La documentation française pour l'utilisateur de l'extension
    \MODULE{paresse} est désormais disponible sous le nom de
    \texttt{paresse-fr.pdf}.

    Le fichier \texttt{paresse.pdf} contient le code commenté en
    français.
  \end{abstract}
\end{otherlanguage}

\tableofcontents{}

\newpage

\section{Introduction}

This package provides only a `quick and low-cost' access to greek
letters which one can obtain with a macro such as \cs{alpha} or
\cs{Omega}. It provides also an environment and a macro which make
possible the use of § to type in those letters. Because of an
\cs{ensuremath} we are not bound to explicitly enter ---i.e. by typing
|$ $| or |\( \)| or else |\[ \]| or anything whatsoever with the same
effect--- mathematics mode to obtain a greek letter.

The idea of the method is from T.~\textsc{Lachand-Robert} and
described in~\cite{tlachand}. I have just add the \cs{ensuremath} which
is so agreeable to write macros.

There is \emph{no} macros for the lowercase omicron nor for the
uppercase alpha, beta\dots{} that one can obtain with the latin roman
letters with the same look. I have not had the courage nor the
strength to build a solution which would provide a means of obtaining
an upright uppercase alpha in a math formula enbedded in an italic
boldfaced text.

Even if the meaning of the French `paresse' is just `lazyness' I would
like to enphasize that the name of this package comes from the fact
that the sign § can be used to point at a paragraph and looks like an
S. So there is no connection between the name and the not unfrequent
sin of the same (French) name\dots{} or maybe\dots{}

\section{Why a 3rd Version?}

With a mail dated march, 3, 2003, Claudio \textsc{Beccari} kindly
informed me that there was an encoding of the greek alphabet with latin
letters some 15~years before I commited this extension. This encoding
was devised by Sylvio \textsc{Levi} who, at the time, was designing
the first greek font for \TeX, using the correspondance between greek
and us keyboard. Claudio wrote to me, and I can't but agree with him,
that if one is used to \textsc{Levi}'s encoding, one would rather keep
one's habit in order to use \MODULE{paresse}.

I, then, decided to provide a new couple of mutually exclusive
options: the first one is \Option{legacy} with which one obtain the
original encoding of this extension and which is active by default,
the other one is \Option{Levi} which provides Sylvio \textsc{Levi}'s
encoding.

I take advantage of this update to make some cosmetic changes: from
now on all inner \emph{secret} macros have a name which begins with
\cs{GA@}; the |.dtx| file is reorganised to facilitate the translation
of the documentation.

\section{Usage}

One loads the package with |\usepackage{paresse}| \textbf{after} the
package |inputenc|. The sign § must be recognised as a letter by
\TeX. On can use for instance |inputenc| with option |latin1| for such
a purpose.

By default the package is loaded with option |wild| and so the macros
such as |§a| are immediately available. If one prefers one can choose
the option \Option{tame} by writing |\usepackage[tame]{paresse}|. One
must then use the command \cs{ActiveLaParesse} or the environment
|ParesseActive| to use the `§-macros'.

When `paresse' is active, one has just to type |§a| in to obtain
\(\alpha\). One has access, by the same means, to all the other greek
letters to which a macro is devoted such as \cs{alpha}, see the
tables~\ref{codageorig} and~\ref{codagelevi}. One obtains
\(\alpha^{\beta}\) with |\(§a^{§b}\)| when § is active. One will note
that, if the package |amsmath| is loaded, the curly braces are
\emph{not} compulsory and that one obtains the same result with just
|\(§a^§b\)|.

\subsection{Options}

\begin{itemize}
\item \Option{tame}\DescribeOption{tame / wild} is the contrary of
  \Option{wild} which is the option by default. When \Option{tame}
  reigns, one \textbf{must} use an environment |ParesseActive| or a
  command \cs{ActiveLaParesse} in order to use the §-macros.

\item \Option{Levi}\NOUVO\DescribeOption{legacy / Levi} is the
  contrary of \Option{legacy} which is the default. With
  \Option{legacy} one uses the original encoding of
  \MODULE{paresse.sty} as it is given by the
  table~\ref{codageorig}. If the option \Option{Levi} is enforced, one
  uses the Sylvio \textsc{Levi} encoding, see the
  table~\ref{codagelevi}.

\item \Option{ttau}\DescribeOption{ttau / ttheta} is the contrary of
  \Option{ttheta} which is selected by default. When \Option{ttheta}
  is active |§t| gives \(\theta\) in the contrary |§t| gives
  \(\tau\). In all cases, \(\theta\) is given by |§v| and \(\tau\) by
  |§y|. \NOUVO That option is ineffective when one has chosen
  \Option{Levi}.

\textbf{Remark:} \NOUVO when one has chosen the option
\Option{legacy}, \(\Theta\) is `regularly' obtained with |§V| and
\emph{also} with |§T| whatever is the chosen option. In the case of
the option \Option{Levi}, |§V| doesn't correspond to any greek letter.

\item \Option{epsilon}\DescribeOption{epsilon / varepsilon} is the
  contrary of \Option{varepsilon} which is selected by default. With
  \Option{epsilon}, |§e| gives \(\epsilon\) otherwide |§e| gives
  \(\varepsilon\).

\item The following `couples' behave as \Option{epsilon},
  \Option{varepsilon}: \Option{theta} and \Option{vartheta};
  \Option{pi} and \Option{varpi}; \Option{rho} and \Option{varrho};
  \Option{sigma} and \Option{varsigma}; \Option{phi} and
  \Option{varphi}.
\end{itemize}

The default options are \Option{varepsilon}, \Option{theta},
\Option{pi}, \Option{rho}, \Option{sigma}, \Option{varphi},
\Option{wild} and \Option{legacy}\NOUVO. That ensures that this 3rd
version behaves, by default, as the preceding one.

\subsection{Commands and Environment}

\DescribeMacro{\makeparesseletter}
This command gives the letter-catcode to the `character' §. After that
one can use § in the name of a macro, for instance. It corresponds to
the well-known |\makeatletter|.

\DescribeMacro{\makeparesseother}
This macro gives the catcode \emph{other} to the character §. It is
the `contrary' of the preceding one. It corresponds to |\makeatother|.

\DescribeMacro{\ActiveLaParesse}
This macro makes § active and thus enable one to access the macros the
name of which begins with § such as |§a|. A list of these macros and
theirs meanings is given in the tables~\ref{codageorig}
and~\ref{codagelevi}.

\DescribeEnv{ParesseActive}
In this environment § is active and one can use the §-macros. One
could use this environment if one want to use the §-macros when the
package \texttt{paresse.sty} is loaded whith the option \texttt{tame}.

\subsection{Tables of the Macros}

\subsubsection{\MODULE{paresse.sty}'s Original Encoding}

This is the active encoding when one choses the option \Option{legacy}
which is the default.

\begin{center}\Large
\begin{ParesseActive} \label{codageorig}
\begin{tabular}{*4{||>{\ttfamily \§}c|c}||}\hline
a & §a & b & §b & g & §g & d & §d\\ \hline
e & §e & z & §z & h & §h & v & §v\\ \hline
i & §i & k & §k & l & §l & m & §m\\ \hline
n & §n & x & §x & p & §p & r & §r\\ \hline
s & §s & y & §y & u & §u & f & §f\\ \hline
c & §c & q & §q & w & §w &\multicolumn{1}{c|}{}&\\ \hline\hline
G & §G & D & §D & V & §V & L & §L\\ \hline
X & §X & P & §P & S & §S & U & §U\\ \hline
F & §F & Q & §Q & W & §W &\multicolumn{1}{c|}{}&\\ \hline
\end{tabular}
\end{ParesseActive}
\end{center}

\paragraph{Remarks : } all the latin letters used in the name of the
§-macros, but for {\ActiveLaParesse §v, §y and §q}, are loaded with
reminiscences, I hope \texttt{:-)} and the greek uppercases are
obtained with the (latin) corresponding uppercases.

\subsubsection{Sylvio \textsc{Levi}'s Encoding}

One make this encoding active with the option \Option{Levi}.

\begin{center}\Large\NOUVO
\begin{ParesseActive} \label{codagelevi}
\begin{tabular}{*4{||>{\ttfamily \§}c|c}||}\hline
a & §a & b & §b & g & §g & d & §d\\ \hline
e & §e & z & §z & h & §h & j & §v\\ \hline
i & §i & k & §k & l & §l & m & §m\\ \hline
n & §n & x & §x & p & §p & r & §r\\ \hline
s & §s & t & §y & u & §u & f & §f\\ \hline
q & §c & y & §q & w & §w & c & \(\varsigma\)\\ \hline\hline
G & §G & D & §D & J & §V & L & §L\\ \hline
X & §X & P & §P & S & §S & U & §U\\ \hline
F & §F & Y & §Q & W & §W &\multicolumn{1}{c|}{}&\\ \hline
\end{tabular}
\end{ParesseActive}
\end{center}

Sylvio \textsc{Levi}'s encoding gives a direct acces to \cs{varsigma}
(\(\varsigma\)) with |§c| and is different from the original encoding
just for the letters {\ActiveLaParesse §v, §y, §c et §q}. Here is a
summary of theses differences:

\begin{center}
\newcommand\CT[1]{\multicolumn{1}{c|}{\texttt{#1}}}
\begin{ParesseActive}
\begin{tabular}{|l|*{7}{c|}}\hline
greek letters
& §v & §y & §c & §q & §V & §Q & \(\varsigma\)\\\hline
original encoding
&\CT{\§v/\§t}&\CT{\§y/\§t}&\CT{\§c}
&\CT{\§q}&\CT{\§V/\§T}&\CT{\§Q}& --- \\\hline
S. \textsc{Levi}'s encoding
&\CT{\§j}&\CT{\§t}&\CT{\§q}&\CT{\§y}
&\CT{\§J}&\CT{\§Y} & \CT{\§c}\\\hline
\end{tabular}
\end{ParesseActive}
\end{center}

\begin{thebibliography}{99}
\addcontentsline{toc}{section}{Bibliographie}
\bibitem{tlachand} T.~\textsc{Lachand-Robert}.
\emph{La maîtrise de \TeX{} et \LaTeX{}}.
Masson, Paris, Milan, Barcelone, \oldstylenums{1995}.\\
\textsc{isbn} : \texttt{2-225-84832-7}.
\end{thebibliography}

\end{document}
\endinput
%%
%% End of file `paresse-en.tex'.
