\documentclass[12pt]{article}
\usepackage[american]{babel} 
\usepackage{csquotes} 
\usepackage[style=mla]{biblatex} 

\usepackage[bookmarks,pdfusetitle]{hyperref}

\bibliography{handbooksamples}


\defbibheading{blank}{}
\renewcommand*{\mlanamedash}{-{}-{}-}

\title{Brief Article}
\author{The Author}

\begin{document}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.1. The Basic Entry: A Book by a Single Author}
\newrefsegment \nocite{fukuyama02aa,berlage94aa,freedman84aa,kurlansky02aa,lecarre01aa,rowling00aa,silver97aa,tatar92aa}

\printbibliography[heading=blank,segment=1]



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.2. An Anthology or a Compilation}
\newrefsegment \nocite{lopate94aa,sevillano86aa,spafford92aa,weisser01aa}

\printbibliography[heading=blank,segment=2]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.3. Two or More Books by the Same Author}
\newrefsegment \nocite{borroff63aa,borroff67aa,borroff79aa,frye57aa,frye62aa,frye91aa,frye57ab}

\printbibliography[heading=blank,segment=3]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.4. A Book by Two or More Authors}
\newrefsegment \nocite{eggins97aa,hutcheon00aa,gilman93aa,gilman93ab,quirk85aa,quirk85ab,marquart94aa,rabkin83aa}

\printbibliography[heading=blank,segment=4]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.5. Two or More Books by the Same Authors}
\newrefsegment \nocite{durant65aa,durant77aa,gilbert95aa,gilbert01aa,gilbert79aa}

\printbibliography[heading=blank,segment=5]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.6. A Book by a Corporate Author}
\newrefsegment 
\nocite{american-medical-association89aa,national-research-council00aa,public-agenda-foundation92aa}

\printbibliography[heading=blank,segment=6]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.7. A Work in an Anthology}
\newrefsegment
\textsc{Note:} Support for indicating reprints is currently limited. If necessary, use the \textbf{addendum} field for a short-term solution.

\nocite{allende92aa,franco98aa,hansberry71aa,hanzlik92aa,more97aa,witchcraft86aa,douglass00aa,franklin92aa}

\printbibliography[heading=blank,segment=7]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.8. An Article in a Reference Book}
\newrefsegment
\nocite{azimuthal93aa,ginsburg02aa,mandarin94aa,mohanty87aa,noon89aa,allen95aa,le-patourel87aa}

\printbibliography[heading=blank,segment=8]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.9. An Introduction, a Preface, a Foreword, or an Afterword}
\newrefsegment
\nocite{borges73aa,coetzee01aa,drabble85aa,hamill94aa,marsalis93aa,sears01aa,brodsky97aa,doody96aa,doctorow82aa}

\printbibliography[heading=blank,segment=9]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.10. Cross-References}
\newrefsegment

\nocite{baker86aa}\nocite{baker93aa}\nocite{oates2000aa}\nocite{lennon86aa}\nocite{angelou86aa}\nocite{lebowitz93aa}\nocite{hurston93aa}\nocite{walker00aa}\nocite{rodriguez00aa}\nocite{kingston00aa}\nocite{agee00aa}\nocite{atwan00aa}

\printbibliography[heading=blank,segment=10]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.11. An Anonymous Book}
\newrefsegment

\nocite{bible84aa}\nocite{virginia93aa}\nocite{NewYorkPublicLibrary97aa}

\printbibliography[heading=blank,segment=11]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.12. An Edition}
\newrefsegment

\nocite{crane75ab}\nocite{crane75aa}\nocite{twain93aa}\nocite{smith02aa}\nocite{shakespeare92aa}\nocite{octovian86aa}\nocite{edgeworth92aa}\nocite{austen01aa}

\printbibliography[heading=blank,segment=12]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.13. A Translation}
\newrefsegment
\textsc{Note:} Biblatex-mla will not allow you to decide the order to display the names for translator and editor.  Compare the following entry on \emph{Beowulf} with that in the \emph{MLA Handbook}.  The deviation is minor.

\nocite{mu76aa}\nocite{hildegard-of-bingen01aa}\nocite{esquivel92aa}\nocite{beowulf01aa}\nocite{murasaki-shikibu76aa}\nocite{levi87aa}\nocite{esquivel92ab}

\printbibliography[heading=blank,segment=13]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.14. A Book Published in a Second or Subsequent Edition}
\newrefsegment

\nocite{murasaki-shikibu85aa}\nocite{hyde86aa}\nocite{cheyfitz97aa}\nocite{chaucer57aa}\nocite{bondanella01aa}

\printbibliography[heading=blank,segment=14]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.15. A Multivolume Work}
\newrefsegment
\textsc{Note:} There are currently some limitations to presenting multivolume works systematically:
\begin{enumerate}
\item There is no test for open ranges in the years to allow for printing of ``vols.\ to date,'' as in ``3 vols.\ to date'' needed in Caro and Cassidy below.
\item There is no capacity to define multiple editors for one work.  If the \textbf{editor} field is defined for a work that is part of a multivolume set, the editor will be printed in a way indicating editorial responsibility for a particular volume.  See, for example, the entry for Crane's \emph{Red Badge of Courage} that indicates the work is the second volume of a 10 volume set; in this example, Fredson Bowers is the editor of all 10 volumes, though he is indicated as the editor of only the one being cited.  The entry for Howells below, too, should indicate the set's general editorship by Edwin H. Cady, though it does not.
\item Finally, the \textbf{origyear} field designates the date of original publication of the single volume except in cases where both \textbf{volume} and \textbf{volumes} are defined---in which instance it designates the years of publication of the volumes.
\end{enumerate}
These limitations can be temporarily overlooked by use of the \textbf{addendum} field to print information at the end of an entry.  In the long term, such a workaround is undesirable because it sacrifices portability of data.

\nocite{howells68aa}\nocite{crane69ab}\nocite{durant65ac}\nocite{churchill57ab}\nocite{caro02ab}\nocite{durant65ab}\nocite{churchill57aa}\nocite{caro02aa}\nocite{wellek86aa}\nocite{stowe02ab}\nocite{doyle93ac}\nocite{stowe02aa}\nocite{lawrence00aa}\nocite{doyle93ab}\nocite{wellek55aa}\nocite{boulton79aa}\nocite{crane69aa}\nocite{churchill56aa}\nocite{cassidy85aa}\nocite{schlesinger73aa}\nocite{sadie80aa}\nocite{rampersand02aa}\nocite{lauter02aa}\nocite{doyle93aa}\nocite{blanco93aa}

\printbibliography[heading=blank,segment=15]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.16. A Book in a Series}
\newrefsegment

\nocite{riccoboni98aa}\nocite{neruda91aa}\nocite{murck00aa}\nocite{mitchell-boyask02aa}

\printbibliography[heading=blank,segment=16]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.17. A Republished Book}
\newrefsegment

\nocite{dreiser00aa}\nocite{douglas93aa}\nocite{doctorow96aa}\nocite{atwood01aa}\nocite{ishiguro01aa}

\printbibliography[heading=blank,segment=17]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.18. A Publisher's Imprint}
\newrefsegment

\nocite{cassidy85aa}\nocite{lopate94aa}\nocite{morrison02aa}

\printbibliography[heading=blank,segment=18]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.19. A Book with Multiple Publishers}
\newrefsegment

\textsc{Note:} Bibtex is unable to handle books with multiple publishers.

%\printbibliography[heading=blank,segment=19]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.20. A Pamphlet}
\newrefsegment

\nocite{washingtondc00aa}\nocite{renoir94aa}

\printbibliography[heading=blank,segment=20]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.21. A Government Publication}
\newrefsegment

\textsc{Note:} Support for government publications is currently limited.

%\printbibliography[heading=blank,segment=21]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.22. The Published Proceedings of a Conference}
\newrefsegment

\nocite{freed91aa}\nocite{hualde00aa}\nocite{chang00aa}

\printbibliography[heading=blank,segment=22]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.23. A Book in a Language Other Than English}
\newrefsegment

\nocite{poche77aa}\nocite{maraini00aa}\nocite{esquivel89aa}\nocite{dahlhaus67aa}\nocite{bessiere90aa}

\printbibliography[heading=blank,segment=23]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.24. A Book Published before 1900}
\newrefsegment

\nocite{segni49aa}\nocite{dewey99aa}\nocite{brome73aa}

\printbibliography[heading=blank,segment=24]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.25. A Book without Stated Publication Information or Pagination}
\newrefsegment
\textsc{Note:} There is an error in the entry for Bauer below.  The question mark in the year field cancels the final period that should appear after the brackets.  I hope to fix this problem in a future version.

\nocite{sendak63aa}\nocite{photographic-viewXXaa}\nocite{michelangelo92aa}\nocite{malachi09aa}\nocite{bauer71aa}

\printbibliography[heading=blank,segment=25]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.26. An Unpublished Dissertation}
\newrefsegment

\nocite{kelly01aa}\nocite{boyle83aa}

\printbibliography[heading=blank,segment=26]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{5.6.27. A Published Dissertation}
\newrefsegment

\nocite{fullerton01aa}\nocite{dietze82aa}

\printbibliography[heading=blank,segment=27]


%\printbibliography
\end{document}  
