% This file is based on Knuth's and Levy's "testfont"files
%
% The text is taken from the book of Kwsto'ula Mhtropo'ulou, Za'ar 19.
%
\tracinglostchars=0
\tolerance=1000
\raggedbottom
\nopagenumbers
\parindent=0pt
\newlinechar=`@
\input rgreekmacros %not necessary if you have a greek format
 
\newcount\m \newcount\n \newcount\p \newdimen\dim
\chardef\other=12
 
\def\today{\ifcase\month\or
  January\or February\or March\or April\or May\or June\or
  July\or August\or September\or October\or November\or December\fi
  \space\number\day, \number\year}
\def\hours{\n=\time \divide\n 60
  \m=-\n \multiply\m 60 \advance\m \time
  \twodigits\n\twodigits\m}
\def\twodigits#1{\ifnum #1\lt10 0\fi \number#1}
 
{\catcode`\|=0 \catcode`\\=\other
|gdef|help{|message{%
\init switches to another font;@%
\end or \bye finishes the run;@%
\table prints the font layout in tabular format;@%
\text prints a sample text, assuming TeX text font conventions;@%
\sample combines \table and \text;@%
\mixture mixes a background character with a series of others;@%
\alternation interleaves a background character with a series;@%
\alphabet prints all lowercase letters within a given background;@%
\alphabets does one \alphabet for each lowercase background letter;@%
\ALPHABET prints all uppercase letters within a given background;@%
\ALPHABETs does one \ALPHABET for each uppercase background letter;@%
\accents prints all lowercase vowels with accents and breathings;@%
\series prints a series of letters withing a given background;@%
\lowers prints a comprehensive test of lowercase;@%
\uppers prints a comprehensive test of uppercase;@%
\digits prints a comprehensive test of numerals;@%
\math prints a comprehensive test of TeX math italic;@%
\names prints a text that mixes upper and lower case;@%
\punct prints a punctuation test;@%
\bigtest combines many of the above routines;@%
\help repeats this message;@%
and you can use ordinary TeX commands (e.g., to \input a file).}}}
 
\def\setbaselineskip{\setbox0=\hbox{\n=0
\loop\char\n \ifnum \n\lt255 \advance\n 1 \repeat}
\baselineskip=6pt \advance\baselineskip\ht0 \advance\baselineskip\dp0 }
 
\def\setchar#1{{\escapechar-1\message{\string#1 character = }%
  \def\do##1{\catcode`##1=\other}\dospecials
  \read-1 to\next
  \expandafter\finsetchar\next\next#1}}
\def\finsetchar#1#2\next#3{\global\chardef#3=`#1
  \ifnum #3=`\# \global\chardef#3=#2 \fi}
\def\promptthree{\setchar\background
  \setchar\starting \setchar\ending}
 
\def\init{\message{@Name of the font to test = }
  \read-1 to\fontname
  \font\testfont=\fontname
  \headline={\sevenrm Test of \fontname\unskip\ on \today\ at \hours\hfil
    Page \folio}
  \testfont \setbaselineskip
  \ifdim\fontdimen6\testfont\lt10pt \rightskip=0pt plus 20pt
  \else\rightskip=0pt plus 2em \fi
  \spaceskip=\fontdimen2\testfont % space between words (\raggedright)
  \xspaceskip=\fontdimen2\testfont \advance\xspaceskip by\fontdimen7\testfont
  \message{Now type a test command (\string\help\space for help):}}
 
\def\mixture{\promptthree \domix\mixpattern}
\def\alternation{\promptthree \domix\altpattern}
\def\mixpattern{\0\1\0\0\1\1\0\0\0\1\1\1\0\1}
\def\altpattern{\0\1\0\1\0\1\0\1\0\1\0\1\0\1\0\1\0}
\def\domix#1{\par\chardef\0=\background \n=\starting
  \until\ifnum\n=\ending\do\chardef\1=\n #1\par\n=\count\n \repeat}
 
\count`w=`a\count`a=`b\count`b=`g\count`g=`d\count`d=`e\count`e=`z
\count`z=`h\count`h=`j\count`j=`i\count`i=`k\count`k=`l\count`l=`m
\count`m=`n\count`n=`x\count`x=`o\count`o=`p\count`p=`r\count`r=`c
\count`c=`t\count`t=`u\count`u=`f\count`f=`q\count`q=`y\count`y=`w
\count`W=`A\count`A=`B\count`B=`G\count`G=`D\count`D=`E\count`E=`Z
\count`Z=`H\count`H=`J\count`J=`I\count`I=`K\count`K=`L\count`L=`M
\count`M=`N\count`N=`X\count`X=`O\count`O=`P\count`P=`R\count`R=`S
\count`S=`T\count`T=`U\count`U=`F\count`F=`Q\count`Q=`Y\count`Y=`W
\count`0=`1\count`1=`2\count`2=`3\count`3=`4\count`4=`5
\count`5=`6\count`6=`7\count`7=`8\count`8=`9\count`9=`0
 
\long\def\until#1\do#2\repeat{\def\body{#2}\def\cond{#1}\etareti}
\def\etareti{\cond\let\next=\relax\else\let\next=\etareti\fi\body\next}
 
\def\alphabet{\setchar\background \dosseries{`a}{`w}}
\def\ALPHABET{\setchar\background \doseries{`A}{`W}}
\def\series{\promptthree \doseries\starting\ending}
\def\doseries#1#2{\def\!{\discretionary{\background}{\background}{\background}}
  \!\n=#1\until\ifnum\n=#2\do\char\n\!\n=\count\n \repeat\par}
\def\dosseries#1#2{\def\!{\discretionary{\background}{\background}{\background}}
  \!\n=#1\until\ifnum\n=#2\do\char\n\!\n=\count\n \repeat s\par}
 
\def\lowers{\docomprehensive hs{`a}{`w}}
\def\uppers{\docomprehensive HO{`A}{`W}}
\def\digits{\docomprehensive H0{`0}{`9}}
\def\docomprehensive#1#2#3#4{\par\n=#3
  \until\ifnum\n=#4\do\line{\altr#1\hfil\altr#2}\n=\count\n\repeat}
\def\altr#1{\setbox0=\hbox{#1}\setbox2=\hbox{#1\char\n}%
  \dim=.5\hsize \advance\dim-\wd0 \divide\dim\wd2 \multiply\dim\wd2
  \xleaders\box2\hskip\dim\box0}
\def\alphabets{\m=`a\until\ifnum\m=`w
  \do\chardef\background=\m{\dosseries{`a}{`w}}\m=\count\m \repeat}
\def\ALPHABETS{\m=`A\until\ifnum\m=`W
  \do\chardef\background=\m{\doseries{`A}{`W}}\m=\count\m \repeat}
 
\def\separator{\vfil\penalty-100\vfilneg
  \medskip\leaders\hrule height.4pt\vskip.4pt\medskip}
 
\def\bigtest{%
\sample\vfil\break
\accents\separator
\names\separator
\punct\separator
\alphabets\separator
\lowers\separator
\uppers\separator
\digits}
 
\def\math{\message{(Sorry, that's not here yet.)}}
 
\def\oct#1{\hbox{\tenrm\'{}\kern-.2em\tenit#1\/\kern.05em}} % octal constant
\def\hex#1{\hbox{\tenrm\H{}\tentt#1}} % hexadecimal constant
\def\setdigs#1"#2{\gdef\h{#2}% \h=hex prefix; \0\1=corresponding octal
 \m=\n \divide\m by 64 \xdef\0{\the\m}%
 \multiply\m by-64 \advance\m by\n \divide\m by 8 \xdef\1{\the\m}}
\def\testrow{\setbox0=\hbox{\penalty 1\def\\{\char"\h}%
 \\0\\1\\2\\3\\4\\5\\6\\7\\8\\9\\A\\B\\C\\D\\E\\F%
 \global\p=\lastpenalty}} % \p=1 if none of the characters exist
\def\oddline{\cr
  \noalign{\nointerlineskip}
  \multispan{19}\hrulefill&
  \setbox0=\hbox{\lower 2.3pt\hbox{\hex{\h x}}}\smash{\box0}\cr
  \noalign{\nointerlineskip}}
\newif\ifskipping
\def\evenline{\loop\skippingfalse
 \ifnum\n\lt128 \m=\n \divide\m 16 \chardef\next=\m
 \expandafter\setdigs\meaning\next \testrow
 \ifnum\p=1 \skippingtrue \fi\fi
 \ifskipping \global\advance\n 16 \repeat
 \ifnum\n=128 \let\next=\endchart\else\let\next=\morechart\fi
 \next}
\def\morechart{\cr\noalign{\hrule}
 \chartline \oddline \m=\1 \advance\m 1 \xdef\1{\the\m}
 \chartline \evenline}
\def\chartline{&\oct{\0\1x}&&\:&&\:&&\:&&\:&&\:&&\:&&\:&&\:&&}
\def\chartstrut{\lower4.5pt\vbox to14pt{}}
\def\table{$$\global\n=0
  \halign to\hsize\bgroup
    \chartstrut##\tabskip0pt plus10pt&
    &\hfil##\hfil&\vrule##\cr
    \lower6.5pt\null
    &&&\oct0&&\oct1&&\oct2&&\oct3&&\oct4&&\oct5&&\oct6&&\oct7&\evenline}
\def\endchart{\cr\noalign{\hrule}
  \raise11.5pt\null&&&\hex 8&&\hex 9&&\hex A&&\hex B&
  &\hex C&&\hex D&&\hex E&&\hex F&\cr\egroup$$\par}
\def\:{\setbox0=\hbox{\char\n}%
  \ifdim\ht0>7.5pt\reposition
  \else\ifdim\dp0>2.5pt\reposition\fi\fi
  \box0\global\advance\n 1 }
\def\reposition{\setbox0=\vbox{\kern2pt\box0}\dim=\dp0
  \advance\dim 2pt \dp0=\dim}
\def\centerlargechars{
  \def\reposition{\setbox0=\hbox{$\vcenter{\kern2pt\box0\kern2pt}$}}}
 
\begingreek
 
\def\accents{\longseries a\shortseries e\longseries h\mediumseries i
\shortseries o\mediumseries u\longseries w \rseries}
\def\longseries#1{#1 `#1 '#1 ~#1 #1| `#1| '#1| ~#1| <#1 <`#1 <'#1 <~#1 <#1|
  <`#1| <'#1| <~#1| >#1 >`#1 >'#1 >~#1 >#1| >`#1| >'#1| >~#1| \par}
\def\shortseries#1{#1 `#1 '#1 <#1 <`#1 <'#1 >#1 >`#1 >'#1 \par}
\def\mediumseries#1{#1 `#1 '#1 ~#1 %
  <#1 <`#1 <'#1 <~#1 >#1 >`#1 >'#1 >~#1 "#1 "`#1 "'#1 "~#1 \par}
\def\rseries{r <r >r\par}
\def\names{ >Afrod'ith Beatr'ikh Gi'annhc Dhm'htrhc E>ugen'ia
   Zw'h <Hrakl~hc Je'ofiloc >Iw'anna Kl'aouc Louk~ac
   Mar'ia N'ikh Xanjo'ula O>id'ipouc Pl'atwn Rox'anh S'ilbio
   Ter'eza <Upat'ia F'ilippoc Qar'a Yuq'h >Wrig'enhc \par}
\def\punct{{\hyphenpenalty=10000\par\dopunct{ka'i}\dopunct{t~hs}\dopunct{<'ena}
  \dopunct{<ELLAS}\par}}
\def\dopunct#1{#1, #1: #1; ''#1'' #1? #1! ((#1)) (#1) [#1] #1* #1.\par}
\def\from#1{{\sevenrm\unskip\nobreak\hfil\penalty50\hskip2em\hbox{}
  \nobreak\hfil (From #1)\parfillskip=0pt\finalhyphendemerits=0 \par}}
 
\def\text{{\advance\baselineskip-4pt
\setbox0=\hbox{abcdefghijklmnopqrstuvwxyz}
\ifdim\hsize\gt2\wd0 \ifdim 15pc\gt2\wd0 \hsize=15pc \else \hsize=2\wd0 \fi\fi
((T'o br~hka!)) f'wnaxe. ((T'o br~hka! E>~isai <'ena pr'oswpo m'esa >ap'o
<'ena bibl'io po'u pot'e d'en gr'afthke, giat'i >eg'w prosp'ajhsa, >eg'w
m'ono, ka'i m''a>ut'h t'h nekr'h gl'wssa ka'i d'en mp'oresa n'a t'o p~w, n'a
t'o fti'axw, d'en mp'oresa pot'e, giat'i <h gl'wssa e>~inai nekr'h,
>endi'amesh, mi'a gl'wssa gi'a poiht'ec, gi'a >anjr'wpouc po'u >oneire'uontai
s'an a>ut'o t'o paid'i, t'on >Ast'erh po'u to~u f'ortwsan >'adika du'o
jan'atouc ka'i d'en m'ilhse. D'en e>~inai mi'a gl'wssa gi'a >emp'orio to'uth
>ed~w <h dik'h mou ka'i gi''a>ut'o d'en s'e >anagn'wrise kan'enac <'otan
milo~usa gi'a s'ena. D'en j'a mporo~use pot'e n'a e>~inai s'ugqronh <h pr'axh
m'e t'h fantas'ia. Pot'e >'idia <h pragmatik'h <istor'ia m'e t'h nekr'h
gl'wssa. Pot'e. T'a pr'oswpa <up'hrxane, na'i, e>~inai >al'hjeia, <up'hrxane
>all'a pot'e s'e sq'esh m'e t'hn pragmatik'othta po'u dihg'hjhka m'e t'h
nekr'h, >'aqrhsth gl'wssa t~wn qam'enwn >Etro'uskwn. <Up'hrxane s'e qr'ono
ka'i q~wro parwqhm'eno, >'asqeto, qr'ono mhd'en, pr'oswpa po'u <h
pragmatik'othta <h dik'h mac t'a e>~iqe de~i >alli~wc. Ka'i m'ono >es'u,
m'ono >es'u e>~isai mis'oc, e>~isai KANENAS, giat'i >eg'w d'en pr'oftasa
n'a s'e k'anw >al'hjeia, >eg'w d'en mp'oresa n'a s'e k'anw pr'oswpo
<uparkt'o m'esa >ap'o t'h fjor'a. E>~isai <'ena pr'oswpo m'esa >ap'o <'ena
mujist'orhma po'u pot'e d'en gr'afthke giat'i <h gl'wssa a>ut'h e>~inai
m'ono gi'a t'oso, t'oso l'igo, <'oso >~htan o<i 19 m'erec s'e ke~ino t'o
Sept'embrh ka'i >'etsi >~htan po'u s''>'afhsa mis'o, s''<'ena trago'udi
m'esa, s'e mi'a l'exh pali'a, s''<'ena >anastenagm'o, s'e mis'o >'oneiro
spasm'eno. S''>'afhsa >'etsi ka'i d'en >'eginec pot'e mi'a >al'hjeia gi'a
to'uc >'allouc, gi'a to'uc tr'itouc, gi'a to'uc logiko'uc, gi'a <'osouc
rwto~usan gi'a s'ena, gi'a t'h guna'ika sou >ak'oma po'u nom'izei p'wc s'e
x'erei ka'i d'en x'erei t'ipota >ap'o s'ena. >Eg'w fta'iw >ag'aph mou ka'i
p'ali >ag'aph j'a so~u''dina ka'i l'exeic nekr'ec, >'aqrhstec ka'i
m''a>ut'ec j'a prospajo~usa, j'a >'hjela n'a g'ineic <'ena >ast'eri ka'i mi'a
ter'astia fwtein'h gramm'h st'hn >'akrh to~u k'osmou, >en~w >es'u mo~u
mil~ac gi'a qrus'afi ka'i p'wc gi''a>ut'o t'o mis'o qrus'afi, mis'o ka'i
klemm'eno, sk'otwsec du'o >aj'wouc, du'o x'enouc, du'o zw'ec, >es'u, t'o
>'astro mou ka'i t'o f~wc mou. M'ilhs'e mou loip'on, p'ec mou p'wc >'eqw
>'adiko ka'i p'wc d'en e>~inai gi'a s'ena <'ola to~uta, p'ec mou t'hn
>al'hjeia >'h <'ena y'ema, p'ec mou, p'ec mou Za'ar >ag'aph mou...))
 
Logik'oc. M'e t'ic l'exeic metrhm'enec. Du'o, t'esseric, >oqt'w, d'wdeka. <O
>'erwt'ac tou kr'athse <'oso krat~ane o<i >epoq'ec. Mi'a >epoq'h. M'e broq'ec
ka'i ke~ino t'o jamp'o k'itrino <'hlio. Sept'embrhc. <'Enac. M'ono. Met'a
xanag'urise st'o sp'iti tou. X'enoc. Ka'i m'onoc. Ka'i >'emeine >eke~i.
X'eqase. >'H o>'ute ki a>ut'o. P'erase st'hn >'allh >'oqjh, st'o pot'ami,
p'anta t'o''lege p'wc <h >'allh >'oqjh e>~inai <h pi'o e>'ukolh. >'Oqi a>ut'h
>ed~w po'u t'hn perpat~ame. <H >'allh. A>ut'h po'u d'en t'hn x'eroume >ak'oma.
Za'ar, t'o d'eqthke n'a t'on l'eei m'ono >eke'inh. Na'i e>~iqe skot'wsei. Du'o
>aj'wouc. >'Enoqouc. D'en e>~iqe shmas'ia. >'Afhse n'a <upoyiasto~un t'on
>Ast'erh. >Eke~inoc >'agnwstoc. Ka'i m'ono gi'a t'o qrus'afi milo~use. <O
>Ast'erhc j'a pono~use. <'Opwc p'onese >eke'inh. >'Oqi, d'en >'hjele n'a
pon~ane o<i >'anjrwpoi. >'Oqi po'u noiaz'otan. >'Oqi po'u d'en t'o >'anteqe.
<'Omwc fob'otan. Fob'otan pol'u k'aje for'a po'u >'emene m'onoc m'e t'on
<eaut'o tou. Ka'i <h logik'h d'en >'eftane. T'ote, d'en >'eftane <h logik'h.
>'Htane gi'a to'uc pol'u >aj'wouc a>ut'h <h logik'h. A>ut'oc tetrag'wnize to'uc
k'uklouc. T'esseric gwn'iec. M'ia, d'uo, tre~ic, t'esseric. Gwn'ies. <'Olec.
Ka'i <h logik'h tou. <O >'erwt'ac tou m'ono mi'a >epoq'h.
 
((>Eg'w d'en >erwte'uomai)) e>~iqe pe~i. D'en t'on e>~iqe pist'eyei. >~Htan <o
k'itrinoc jamp'oc Sept'embrhc ka'i <h l'imnh ka'i o<i stajmo'i ka'i o<i n'uqtec
m'e t'on >'erwta. D'en t'on e>~iqe pist'eyei.
 
((T'i >'eqeic n'a mo~u pe~ic t'wra? S''t'o e>~iqa pe~i. <H gl'wssa po'u
>epim'eneic n'a mil~ac, <h dik'h sou <h gl'wssa e>~inai >'aqrhsth >entel~wc
ka'i d'en mpore~i kan'enac n'a t'hn zwntan'eyei. >Eg'w >'emeina p'anta <o
x'enoc, t'o e>~idec, t'o x'ereic, t'o''majec pi'a. Giat'i >epim'eneic? T'i
zht~ac >ap'o m'ena? >Eg'w >'etsi <'opwc e>~imai ftiagm'enoc d'en qwr'aw s'e
kan'ena >'oneiro, d'en qwr'aw s'e kami'a nekr'h di'alekto. E>~imai <o x'enoc
ka'i d'en j'a s''>anagnwr'izw s'e l'igo. D'en j'a x'erw giat'i ka'i p~wc ka'i
p'oso m''>ag'aphsec. Ka'i j'a''jela st''>al'hjeia n'a mo~u t'o pe~ic a>ut'o:
Nom'izeic <'oti m''>ag'amhsec? >Es'u, >em'ena!! M'a poi'a sq'esh j'a mporo~use
n'a <up'arqei >an'ames'a mac? M'ila mou. S'>ako'uw. E>~imai <o x'enoc. >'Emeina
<o x'enoc. K'anenac d'en kat'alabe t'hn >erhmi'a mou. Pot'e. M'onoc ka'i
x'enoc. Ka'i s'e koit'azw st'a m'atia. P~wc s'e l'ene? Poi'a e>~isai? Giat'i
s''>ar'esei n'a >epim'eneic <'oti m''>ag'aphsec? T'i x'ereic >es'u >ap'o
m'ena?))}}
 
\def\sample{\names\table\text}
 
\init
 
