\documentstyle[lingmacros]{article}

\title{Linguistic Macros}
\author{Emma Pease}

\begin{document}

\maketitle

The lingmacros file contains a few macros that may be of use to
linguists.  The major commands are
\begin{enumerate}
\item The enumsentence macros for creating example sentences
\begin{center}
\begin{verbatim}
\enumsentence[label]{sentence}
\eenumsentence[label]{\item[label] sentence 1
                      \item[label] sentence 2}
\end{verbatim}
\end{center}
 The enumsentence macros are similar to the math equation environment
except that they take regular text.  The \verb+\label+ command can be
used inside of them.\footnote{The enumsentence macros use the counter,
enums.  For more information read the style file.} Inside of the
eenumsentence either the \verb+\toplabel+ command for the sentence
number only or the \verb+\label+ for sentence number and sublabel.
Also defined is
\verb+\ex{number}+ which gives a relative reference.  \verb+(\ex{1})+
or \verb+(\ref{senta})+ give the number of the next enumsentence (\ex{1})
or (\ref{senta}); \verb+(\ex{2})+ or \verb+(\ref{sentb})+ of the second
succeeding enumsentence (\ex{2}) or (\ref{sentb}).
\verb+(\ref{sentbb})+ will give (\ref{sentbb}).
 \enumsentence{This is the first sentence\label{senta}}
\eenumsentence{\item this is a second sentence \toplabel{sentb}
               \item this is a third sentence\label{sentbb}}
\begin{center}
\begin{verbatim}
\enumsentence{This is the first sentence\label{senta}}
\eenumsentence{\item this is a second sentence \toplabel{sentb}
               \item this is a third sentence\label{sentbb}}
\end{verbatim}
\end{center}


\item The tree macros
\begin{center}
\begin{verbatim}
\smalltree{alignment structure}
\modsmalltree{number of columns}{alignment structure}
\end{verbatim}
\end{center}
\verb+\smalltree+ is based on the tabular environment with a large
baselineskip. A simple example should suffice.
\enumsentence{\smalltree{& &a\\ 
                         &b& &c\\ 
                        d& &e& &f}}
\begin{center}
\begin{verbatim}
\enumsentence{\smalltree{& &a\\ 
                         &b& &c\\ 
                        d& &e& &f}}
\end{verbatim}
\end{center}
\verb+\modsmalltree+ sometime works better.
\enumsentence{\modsmalltree{5}{& &a\\ 
                         &b& \mc{3}{this is long}\\ 
                        d& &e& &f}}
\begin{center}
\begin{verbatim}
\enumsentence{\modsmalltree{5}{& &a\\ 
                         &b& \mc{3}{this is long}\\ 
                        d& &e& &f}}
\end{verbatim}
\end{center}
 Note the use of the \verb+\mc{number of columns}{text}+ to span
several columns.  The \verb+\clap{text}+ might also be useful; it is
similar to the \verb+\rlap+ and \verb+\llap+ commands and produces a
centered hbox of zero width.  Lines need to be drawn in by hand or one
can use the tree-dvips macros. 

\item The gloss macros
\begin{center}
\begin{verbatim}
\shortex{number of columns}{first line}{second line}{gloss}
\shortexnt{number of columns}{first line}{second line}
\end{verbatim}
\end{center}
These can be combined to produce most of the glosses that linguists
should need.
\enumsentence{\shortex{6}{Was & ist & dem & Kind & geschenkt&worden?}
                         {What& is  & the & child& given    &been?}
                         {What has been given to the child?}
               \item \shortexnt{7}
{Das & Fin\'anzamt      & hat & ihn & geschnappt &(und & nicht}
{the &finance authority & has & him & caught     &(and &not}

\shortex{2}{die &Polizei).}
           {the &police).}
           {It was the IRS that caught him (and not the police).}}
\begin{center}
\begin{verbatim}
\enumsentence{\shortex{6}{Was & ist & dem & Kind & geschenkt&worden?}
                         {What& is  & the & child& given    &been?}
                         {What has been given to the child?}}
               \item \shortexnt{7}
{Das & Fin\'anzamt      & hat & ihn & geschnappt &(und & nicht}
{the &finance authority & has & him & caught     &(and &not}

\shortex{2}{die &Polizei).}
           {the &police)}
           {It was the IRS that caught him (and not the police).}}
\end{verbatim}
\end{center}
Unfortunately, I've not figured out a fullproof method of breaking
the glosses automatically so they have to be done by hand.  

\item AVM structures
\begin{center}
\begin{verbatim}
\outerfs{text}
\end{verbatim}
\end{center}
\enumsentence{\evnup[-3pt]{\outerfs{SUBJ & \outerfs{POSS & [PRED `Louise']\\
                                             PRED &`mother$\langle$Poss$\rangle$}\\[3ex]
                             OBJ  & \outerfs{PRED & `PRO'\\
                                             REFL & +}\\
                             PRED & `dress$\langle$SUBJ OBJ$\rangle$'}}}
\begin{center}
\begin{verbatim}
\enumsentence{\evnup[-3pt]{%
\outerfs{SUBJ & \outerfs{POSS & [PRED `Louise']\\
                         PRED &`mother$\langle$Poss$\rangle$}\\[3ex]
         OBJ  & \outerfs{PRED & `PRO'\\
                        REFL & +}\\
         PRED & `dress$\langle$SUBJ OBJ$\rangle$'}}}
\end{verbatim}
\end{center}
\end{enumerate}
\end{document}